Tools for reproducible ex-post survey data harmonization

### 2020-07-22 data-raw/article.Rmd
first few paragraphs - technology overview and grammar

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::write_bib(c(
  .packages(), 'haven', 'labelled', 'vctrs'
), 'packages.bib')
```

@@

The increasing availability of microdata creates new opportunities for joint exploitation of these resources for comparative research. Before joint analysis is possible, the data need to be harmonized, i.e. transformed and combined in a single data set in a way that matches the measurement of the same characteristics and standardizes the coding of responses. We introduce the `retroharmonize` and `eurobarometer` packages that help the reproducible ex-post harmonization of survey data stored in various, individual files with rich metadata.  We illustrate the proposed workflow using the case of the Eurobarometer surveys, which provide rich data on European publics in form of single-wave SPSS files from a period of over 25 years. Our tools are flexible and can be applied to other survey projects.

@@

* Define a common [taxonomy](#taxonomy), probably based on a thesaurus for consistentcy.

@@

--	Creating a [joint data table](#datajoin) where each variable that represents the same class and metadata attributes; documenting all the changes and 	Providing methods that re-export the data files to SPSS, Stata, SAS.

@@

Our workflow generally fits into the tidyverse.

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::write_bib(c(
  .packages(), 'haven', 'labelled', 'vctrs', 'retroharmonize'
), 'packages.bib')
```

@@

The increasing availability of microdata creates new opportunities for joint exploitation of these resources for comparative research. Before joint analysis is possible, the data needs to be harmonized. This means a transformation and combination in a way that the same measurements are coded in the same way.

We introduce the `retroharmonize` and `eurobarometer` packages that help the reproducible, metadata-rich ex-post harmonization of survey data stored in various individual files.  We illustrate the proposed workflow using the case of the Eurobarometer surveys, which provide rich data on European publics in form of single-wave SPSS files from a period of over 25 years. Our tools are flexible and can be applied to other survey projects: we also introduce [a small case study](http://retroharmonize.satellitereport.com/articles/afrobarometer.html) with Afrobarometer.

@@

* Define a common [taxonomy](#taxonomy) based on a thesaurus for consistency.

@@

--	Creating a [joint data table](#datajoin) where each variable that represents the same class and metadata attributes; documenting all the changes and providing methods that re-export the data files to SPSS, Stata, SAS.

@@

The workflow that we created fits into the [tidyverse](https://www.tidyverse.org/), and relies on [dplyr](https://dplyr.tidyverse.org/), [tidyr](https://tidyr.tidyverse.org/) for data processing, [haven](https://haven.tidyverse.org/) for importing,  [tibble](https://tibble.tidyverse.org/) for more modern data framing.

For easier further development, we rely on the r-lib family of [rlang](https://rlang.r-lib.org/) for easier exception handling, [vctrs](https://vctrs.r-lib.org/) for type casting and coercion, and implementation of class methods, and [fs](https://fs.r-lib.org/) for platform-independent file management. Of course, the tidyverse relies on these modern re-implementations of base R and they improve the tidy workflow and benefit from the efficiency gains with of a basic C++ infrastructure.
